An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-03-18 EconGrowthSeminar.Rnw
Now using \newcommand for repeated text in the presentation, such as production functions and rows in variable tables.

@@

\begin{frame}{No Shortage of Economic Growth Models} %----------------------------
  Single Factor: $y = \mathrm{e}^{\lambda (t-t_0)}f^{m}$
  \vspace{\baselineskip}
  \pause

  Cobb-Douglas: $y = \mathrm{e}^{\lambda (t-t_0)}k^{\alpha}l^{\beta}$
  \vspace{\baselineskip}
  \pause

  Cobb-Douglas with Energy: $y = \mathrm{e}^{\lambda (t-t_0)}k^{\alpha}l^{\beta}e^{\gamma}$
  \vspace{\baselineskip}
  \pause

  Constant Elasticity of Substitution:
  \small
  $y = \gamma \mathrm{e}^{\lambda(t-t_0)} \left\{\delta \left[\delta_1 k^{-\rho_1} + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} + (1-\delta) e^{-\rho} \right\}^{-1/\rho}$
  \normalsize
  \vspace{\baselineskip}
  \pause

  LINEX: $y = e \: \mathrm{e}^{\left[a_0 \left(2 - \frac{l + e}{k} \right) + a_0 c_t \left(\frac{l}{e} - 1 \right)\right]}$
  \vspace{\baselineskip}
  \pause

@@

\begin{frame}{Single Factor} %----------------------------
  \begin{equation*} \label{eq:Single_Factor_Generic}
    y = \mathrm{e}^{\lambda (t-t_0)}f^{m}
  \end{equation*}
  \begin{center}
    \begin{tabular}{|c|l|}\hline
      $y$   & $Y$/$Y_0$ \\
      $t$   & year\\
      $t_0$ & initial year\\
      $f$   & $F$/$F_0$, a factor of production\\
      \hline
      $\lambda$   &   technological progress\\
      $m$   &         ``strength'' of factor $f$\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}

\begin{frame}{\includegraphics[height=0.25in]{figure_other/flag_us.pdf} Single-Factor Fit} %----------------------------
<<Results_Table_US_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
sfUSK <- singleFactorModel("US", "K")
sfUSL <- singleFactorModel("US", "L")
sfUSQ <- singleFactorModel("US", "Q")
sfUSX <- singleFactorModel("US", "X")
sfUSU <- singleFactorModel("US", "U")
@
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        $y = \mathrm{e}^{\lambda (t-t_0)}f^{m}$
        \vspace{\baselineskip}

@@

\begin{frame}{\includegraphics[height=0.25in]{figure_other/flag_za.png} Single-Factor Fit} %----------------------------
<<Results_Table_ZA_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
sfZAK <- singleFactorModel("ZA", "K")
sfZAL <- singleFactorModel("ZA", "L")
sfZAQ <- singleFactorModel("ZA", "Q")
sfZAX <- singleFactorModel("ZA", "X")
sfZAU <- singleFactorModel("ZA", "U")
@
  \begin{columns}[c]
    \column{2.2in}
%       \begin{center}
%         $y = \mathrm{e}^{\lambda (t-t_0)}f^{m}$
%         \vspace{\baselineskip}
%         
%         \tiny
%           \begin{tabular}{|c|c|c|}\hline
%             $f$   & $\lambda$                      & $m$\\
%             \hline
%             $k$   & \Sexpr{coef(sfZAK)["lambda"]}  & \Sexpr{coef(sfZAK)["m"]}\\
%             \hline
%             $l$   & \Sexpr{coef(sfZAL)["lambda"]}  & \Sexpr{coef(sfZAL)["m"]}\\
%             \hline
%             $q$   & \Sexpr{coef(sfZAQ)["lambda"]}  & \Sexpr{coef(sfZAQ)["m"]}\\
%             \hline
%             $x$   & \Sexpr{coef(sfZAX)["lambda"]}  & \Sexpr{coef(sfZAX)["m"]}\\
%             \hline
%             $u$   & \Sexpr{coef(sfZAU)["lambda"]}  & \Sexpr{coef(sfZAU)["m"]}\\
%             \hline
%           \end{tabular}
%         \normalsize
%       \end{center}
    \column{2.2in}
<<Results_Graph_ZA_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createSFLatticeGraph("ZA", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
\end{frame}

\begin{frame}{Cobb-Douglas} %----------------------------
  \begin{equation*} \label{eq:CD_Without_Energy}
    y = \mathrm{e}^{\lambda (t-t_0)}k^{\alpha}l^{\beta}
  \end{equation*}
  \begin{equation*} \label{eq:CD_Const_Returns_To_Scale}
    \alpha + \beta = 1
  \end{equation*}
  \begin{equation*} \label{eq:CD_Constraints}
    0 \leq \alpha, \beta \leq 1
  \end{equation*}
  \begin{center}
    \begin{tabular}{|c|l|}\hline
      $k$   & $K$/$K_0$, capital stock\\
      $l$   & $L$/$L_0$, labor\\
      \hline
      $\lambda$   &   technological progress\\
      $\alpha$   &    ``strength'' of capital\\
      $\beta$   &    ``strength'' of labor\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}

@@

\begin{frame}{Cobb-Douglas With Energy} %----------------------------
  \begin{equation*} \label{eq:CDe_With_Energy}
    y = \mathrm{e}^{\lambda (t-t_0)}k^{\alpha}l^{\beta}e^{\gamma}
  \end{equation*}
  \begin{equation*} \label{eq:CDe_Const_Returns_To_Scale}
    \alpha + \beta + \gamma = 1
  \end{equation*}
  \begin{equation*} \label{eq:CDe_Constraints}
    0 \leq \alpha, \beta, \gamma \leq 1
  \end{equation*}
  \begin{center}
    \begin{tabular}{|c|l|}\hline
      $k$   & $K$/$K_0$, capital stock\\
      $l$   & $L$/$L_0$, labor\\
      $e$   & $E$/$E_0$, any of $q$, $x$, or $u$\\
      \hline
      $\lambda$   &   technological progress\\
      $\alpha$   &    ``strength'' of capital\\
      $\beta$   &    ``strength'' of labor\\
      $\gamma$   &    ``strength'' of energy\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}

\begin{frame}{Parameter Estimation for Cobb-Douglas With Energy} %----------------------------
  \begin{equation*}
    y = \mathrm{e}^{\lambda (t-t_0)}k^{\alpha}l^{\beta}e^{\gamma}
  \end{equation*}
  \begin{itemize}
    \item{Objective: determine $\lambda$, $\alpha$, $\beta$, and $\gamma$ that best fits historical data}
    \item{Almost identical to Cobb-Douglas without energy}
    \item{Perform grid search over $\gamma$}
    \item{Re-estimate $\lambda$, $\alpha$, and $\beta$ at each value of $\gamma$}
    \item{Respect sum constraint ($\alpha + \beta + \gamma = 1$)}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Constant Elasticity of Substitution (CES)} %----------------------------
  \begin{equation*} \label{eq:micEconCES_model}
    \begin{split}
    y = & \gamma \mathrm{e}^{\lambda(t-t_0)} \\
        & \cdot \left\{\delta \left[\delta_1 k^{-\rho_1} + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
                        + (1-\delta) e^{-\rho} \right\}^{-1/\rho}
    \end{split}
  \end{equation*}
  \begin{center}
    \begin{tabular}{|c|l|}\hline
      $k$   & $K$/$K_0$, capital stock\\
      $l$   & $L$/$L_0$, labor\\
      $e$   & $E$/$E_0$, any of $q$, $x$, or $u$\\
      \hline
      $\gamma$   &   fitting coefficient\\
      $\lambda$  &   technological progress\\
      $\delta_1$   &   split between ($k$) and ($l$)\\
      $\rho_1$     &   ``strength'' of ($k$) vs. ($l$)\\
      $\delta$   &   split between ($kl$) and ($e$)\\
      $\rho$     &   ``strength'' of ($kl$) vs. ($e$)\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}

@@

\begin{frame}{LINEX} %----------------------------
  \begin{equation*} \label{eq:LINEX_model}
    y = e \: \mathrm{e}^{\left[a_0 \left(2 - \frac{l + e}{k} \right) + a_0 c_t \left(\frac{l}{e} - 1 \right)\right]}
  \end{equation*}
  \begin{center}
    \begin{tabular}{|c|l|}\hline
      $k$   & $K$/$K_0$, capital stock\\
      $l$   & $L$/$L_0$, labor\\
      $e$   & $E$/$E_0$, any of $q$, $x$, or $u$\\
      \hline
      $a_0$   &   fitting parameter\\
      $c_t$   &   fitting parameter\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}

>>>>

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
% A bunch of useful commands for this presentation
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

% The equation for a single-factor model
\newcommand{\sfEqn}{$y = \mathrm{e}^{\lambda (t-t_0)}f^{m}$}
% The equation for a cobb-douglas without energy model
\newcommand{\cdEqn}{$y = \mathrm{e}^{\lambda (t-t_0)}k^{\alpha}l^{\beta}$}
% The equation for a cobb-douglas with energy model
\newcommand{\cdeEqn}{$y = \mathrm{e}^{\lambda (t-t_0)}k^{\alpha}l^{\beta}e^{\gamma}$}
% The equation for a ces model
\newcommand{\cesEqn}{$y = \gamma \mathrm{e}^{\lambda(t-t_0)} \left\{\delta \left[\delta_1 k^{-\rho_1} + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} + (1-\delta) e^{-\rho} \right\}^{-1/\rho}$}
% The equation for a linex model
\newcommand{\linexEqn}{$y = e \: \mathrm{e}^{\left[a_0 \left(2 - \frac{l + e}{k} \right) + a_0 c_t \left(\frac{l}{e} - 1 \right)\right]}$}

% Some commands for parameter tables in slides
\newcommand{\yRow}{$y$ & $Y/Y_0$, output}
\newcommand{\timeRow}{$t$ & year}
\newcommand{\initialTimeRow}{$t_0$ & initial year}
\newcommand{\kRow}{$k$ & $K/K_0$, capital stock}
\newcommand{\lRow}{$l$ & $L/L_0$, labor}
\newcommand{\eRow}{$e$ & $E/E_0$, any of $q$, $x$, or $u$}
\newcommand{\lambdaRow}{$\lambda$ & technological progress}
\newcommand{\alphaRow}{$\alpha$ & ``strength'' of capital}
\newcommand{\betaRow}{$\beta$ & ``strength'' of labor}
\newcommand{\gammaRow}{$\gamma$ & ``strength'' of energy}

@@

\begin{frame}{No Shortage of Economic Growth Models} %----------------------------
  Single Factor: \sfEqn
  \vspace{\baselineskip}
  \pause

  Cobb-Douglas: \cdEqn
  \vspace{\baselineskip}
  \pause

  Cobb-Douglas with Energy: \cdeEqn
  \vspace{\baselineskip}
  \pause

  Constant Elasticity of Substitution:
  \small
  \cesEqn
  \normalsize
  \vspace{\baselineskip}
  \pause

  LINEX: \linexEqn
  \vspace{\baselineskip}
  \pause

@@

\begin{frame}{Single Factor} %----------------------------
  \begin{center}
    \sfEqn
    \vspace{\baselineskip}

    \begin{tabular}{|c|l|}\hline
      \yRow\\
      \timeRow\\
      \initialTimeRow\\
      $f$ & $F$/$F_0$, a factor of production\\
      \hline
      \lambdaRow\\
      $m$ & ``strength'' of factor $f$\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}

\begin{frame}{\includegraphics[height=0.25in]{figure_other/flag_us.pdf} Single-Factor Fit} %----------------------------
<<Results_Table_US_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
sfUSK <- singleFactorModel("US", "K")
sfUSL <- singleFactorModel("US", "L")
sfUSQ <- singleFactorModel("US", "Q")
sfUSX <- singleFactorModel("US", "X")
sfUSU <- singleFactorModel("US", "U")
@
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \sfEqn
        \vspace{\baselineskip}

@@

% \begin{frame}{\includegraphics[height=0.25in]{figure_other/flag_za.png} Single-Factor Fit} %----------------------------
% <<Results_Table_ZA_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
% sfZAK <- singleFactorModel("ZA", "K")
% sfZAL <- singleFactorModel("ZA", "L")
% sfZAQ <- singleFactorModel("ZA", "Q")
% sfZAX <- singleFactorModel("ZA", "X")
% sfZAU <- singleFactorModel("ZA", "U")
% @
%   \begin{columns}[c]
%     \column{2.2in}
%       \begin{center}
%         $y = \mathrm{e}^{\lambda (t-t_0)}f^{m}$
%         \vspace{\baselineskip}
%         
%         \tiny
%           \begin{tabular}{|c|c|c|}\hline
%             $f$   & $\lambda$                      & $m$\\
%             \hline
%             $k$   & \Sexpr{coef(sfZAK)["lambda"]}  & \Sexpr{coef(sfZAK)["m"]}\\
%             \hline
%             $l$   & \Sexpr{coef(sfZAL)["lambda"]}  & \Sexpr{coef(sfZAL)["m"]}\\
%             \hline
%             $q$   & \Sexpr{coef(sfZAQ)["lambda"]}  & \Sexpr{coef(sfZAQ)["m"]}\\
%             \hline
%             $x$   & \Sexpr{coef(sfZAX)["lambda"]}  & \Sexpr{coef(sfZAX)["m"]}\\
%             \hline
%             $u$   & \Sexpr{coef(sfZAU)["lambda"]}  & \Sexpr{coef(sfZAU)["m"]}\\
%             \hline
%           \end{tabular}
%         \normalsize
%       \end{center}
%     \column{2.2in}
% <<Results_Graph_ZA_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
% createSFLatticeGraph("ZA", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
% @
%   \end{columns}
% \end{frame}

\begin{frame}{Cobb-Douglas} %----------------------------
  \begin{center}
    \cdEqn\\
    $\alpha + \beta = 1$\\
    $0 \leq \alpha, \beta \leq 1$
    \vspace{\baselineskip}

    \begin{tabular}{|c|l|}\hline
      \yRow\\
      \timeRow\\
      \initialTimeRow\\
      \kRow\\
      \lRow\\
      \hline
      \lambdaRow\\
      \alphaRow\\
      \betaRow\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}

@@

\begin{frame}{Cobb-Douglas With Energy} %----------------------------
  \begin{center}
    \cdeEqn\\
    $\alpha + \beta + \gamma = 1$\\
    $0 \leq \alpha, \beta, \gamma \leq 1$
    \vspace{\baselineskip}

    \begin{tabular}{|c|l|}\hline
      \yRow\\
      \timeRow\\
      \initialTimeRow\\
      \kRow\\
      \lRow\\
      \eRow\\
      \hline
      \lambdaRow\\
      \alphaRow\\
      \betaRow\\
      \gammaRow\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}

\begin{frame}{Parameter Estimation for Cobb-Douglas With Energy} %----------------------------
  \begin{center}
    \cdeEqn
  \end{center}
  \begin{itemize}
    \item{Objective: determine $\lambda$, $\alpha$, $\beta$, and $\gamma$ that best fits historical data}
    \item{Almost identical to Cobb-Douglas without energy}
    \item{Perform grid search over $\gamma$}
    \item{Re-estimate $\lambda$, $\alpha$, and $\beta$ at each value of $\gamma$}
    \item{Respect sum constraint ($\alpha + \beta + \gamma = 1$)}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Constant Elasticity of Substitution (CES)} %----------------------------
  \begin{center}
    \small
      \cesEqn
    \normalsize
    \vspace{\baselineskip}

    \begin{tabular}{|c|l|}\hline
      \yRow\\
      \timeRow\\
      \initialTimeRow\\
      \kRow\\
      \lRow\\
      \eRow\\
      \hline
      $\gamma$   &   fitting coefficient\\
      \lambdaRow\\
      $\delta_1$   &   split between ($k$) and ($l$)\\
      $\rho_1$     &   ``strength'' of ($k$) vs. ($l$)\\
      $\delta$   &   split between ($kl$) and ($e$)\\
      $\rho$     &   ``strength'' of ($kl$) vs. ($e$)\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}

@@

\begin{frame}{LINEX} %----------------------------
  \begin{center}
    \linexEqn
    \vspace{\baselineskip}

    \begin{tabular}{|c|l|}\hline
      \kRow\\
      \lRow\\
      \eRow\\
      \hline
      $a_0$ & fitting parameter\\
      $c_t$ & fitting parameter\\
      \hline
    \end{tabular}
  \end{center}
\end{frame}
